### 2007-10-08 pkg/paper/article.Rnw
some enhancements/comments in gradient plot
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/colorspace@8 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

<<echo=FALSE>>=
library(grid)
library(colorspace)

@@

# RGB
N <- 100
grid.rect(x=1:N/N, width=1/N, y=.75, height=0.4,
          just=c("right", "centre"),
          gp=gpar(col=NA,
            fill=rgb(seq(165, 255, length=N),
              seq(42, 255, length=N),
              seq(42, 255, length=N), max=255)))

# HCL

# col2rgb("brown")
#       [,1]
# red    165
# green   42
# blue    42
# N <- 20

# > as(RGB(165/255, 42/255, 42/255), "polarLAB")
#             L        C        H
# [1,] 58.72249 41.92013 24.42661

cols <- coords(as(polarLAB(seq(58.7224, 99, length=N),
                           seq(41.92013, 0, length=N),
                           24.42661), "RGB"))

grid.rect(x=1:N/N, width=1/N, y=0.25, height=0.4,
          just=c("right", "centre"),
          gp=gpar(col=NA, fill=rgb(cols[,1], cols[,2], cols[,3])))

@@

The color of light is determined by its wave-length. A typical human can 
perceive light in the range 400nm to 700nm.  A coherent
light source (e.g., a laser) at the 400nm range appears blue and a coherent
light source at the 700nm range appears red.  Most light sources
(e.g., light bulbs, or the Sun)
are not coherent and emit light at many different wave lengths.
A light source can be characterised by its Spectral Power Distribution (SPD),
which is how much of each wavelength of light the light source emits.

@@

\subsection{Device-independence:  CIE XYZ}

Based on experiments with human subjects, the CIE developed the
CIE XYZ colorspace.  In this colorspace, 
any SPD, that is any color, can be mapped to a unique triplet
of XYZ values, which represent the standard observer's perception of
the color.

@@

This is a good first step, but CIE XYZ does not give us 
everything we need.  For example, if we want to specify a
simple color, such as ``red'', which three values should we use?

In this colorspace, $Y$ corresponds roughly to how bright a color
appears, but the $X$ and $Z$ components have no simple intuitive 
meaning.  Furthermore, unit steps within this colorspace are
not perceptually equal, so generating a palette of colors cannot
proceed in a rational manner.  This leads us on to consideration 
to other colorspaces that are derived from CIE XYZ.

\subsection{Perceptual uniformity:  CIE LUV}

The CIE LUV colorspace
 is a transformation of CIE XYZ in a 
reasonably successful attempt to achieve perceptual uniformity.
Again, any color can be specified as a triplet of values,
with the advantage that a unit step in any direction corresponds to a 
fixed perceptual change in color.  However, the problem of 
intuitively specifying colors remains.

\subsection{Intuitive color choice:  CIE LCH}

The CIE LCH colorspace is a re-expression of CIE LUV using
polar coordinates for the U-V plane.  In this colorspace,
any color may be described by a triplet of values, a unit
step is perceptually uniform throughout the colorspace, 
and the three dimensions of the colorspace have intuitive 
interpretations:

@@

Only one problem remains:  we can conveniently specify a color
using CIE LCH, but in order to render that color we need a
specification that a rendering device understands.
For example, in order to display a color on screen,
we need to be able to convert a CIE LCH color into an RGB
color specification.

@@

The conversion from CIE LCH to RGB consists of a series
of steps.  We start with conversions between the various
CIE colorspaces.

@@

This conversion is non-linear, reflecting the non-linear
nature of human color perception.  It is also performed in 
several stages, reflecting the historical development of the
CIE LUV colorspace, which builds upon a couple of other
earlier transformations of the CIE XYZ colorspace.

The conversion also depends on the concept of a \emph{white point}.
For now, we can just think of the white point as representing the
viewing conditions under which the colour is to be shown.
\footnote{It is relatively straightforward 
to think about a white point in terms of a printed image.  In that case,
the white point is an area of the image that perfectly reflects the
ambient lighting conditions.  In other words, the white point
corresponds to whatever illumination is being used to view the image.}
The white point is important because the perception of colors
is dependent on the viewing conditions.  For example, an image
will look different when viewed under a pink light instead
of a white light.  More subtly, 
an image will look different when viewed outside in normal sunlight compared
to inside with fluorescent lighting.

The white point is expressed as CIE XYZ tristimulus values.
There are a number of established white points used in practice,
with the most popular being a representation of normal daylight,
called $D65$:

@@

One such standard is the ITU-R BT.709-2 Recommendation (for
high-definition television), which specifies CIE XYZ values
for the $R$, $G$, and $B$ primaries of a display device.

>>>>

<<echo=FALSE>>=
library("grid")
library("colorspace")
@

@@

# RGB
N <- 100

col <- "brown"
col.rgb <- as.vector(col2rgb(col))
col.lch <- as.vector(coords(as(RGB(col.rgb[1]/255, col.rgb[2]/255, col.rgb[3]/255), "polarLAB")))

fill.rgb <- rgb(seq(col.rgb[1], 255, length=N),
                seq(col.rgb[2], 255, length=N),
                seq(col.rgb[3], 255, length=N), max=255)
fill.lch <- coords(as(polarLAB(seq(col.lch[1], 99, length=N),
                               seq(col.lch[2], 0, length=N),
                                   col.lch[3]), "RGB"))
fill.lch <- rgb(fill.lch[,1], fill.lch[,2], fill.lch[,3]) ## we do not even get this behaviour with
                                                          ## hex(..., gamma = 1)
							  ## only for the inverse there is
							  ## hex2RGB(..., gamma = NA)

grid.rect(x=1:N/N, width=1/N, y=.75, height=0.4,
          just=c("right", "centre"),
          gp=gpar(col=NA,
          fill=fill.rgb))

grid.rect(x=1:N/N, width=1/N, y=0.25, height=0.4,
          just=c("right", "centre"),
          gp=gpar(col=NA,
          fill=fill.lch))

@@

The color of light is determined by its wave-length. A typical human can 
perceive light in the range 400nm to 700nm.  A coherent
light source (e.g., a laser) at the 400nm range appears blue and a coherent
light source at the 700nm range appears red.  Most light sources
(e.g., light bulbs, or the Sun)
are not coherent and emit light at many different wave lengths.
A light source can be characterised by its spectral power distribution (SPD),
which is how much of each wavelength of light the light source emits.

@@

\subsection{Device-independence:  CIEXYZ}

Based on experiments with human subjects, the CIE developed the
CIEXYZ colorspace.  In this colorspace, 
any SPD, that is any color, can be mapped to a unique triplet
of XYZ values, which represent the standard observer's perception of
the color.

@@

This is a good first step, but CIEXYZ does not give us 
everything we need.  For example, if we want to specify a
simple color, such as ``red'', which three values should we use?

In this colorspace, $Y$ corresponds roughly to how bright a color
appears, but the $X$ and $Z$ components have no simple intuitive 
meaning.  Furthermore, unit steps within this colorspace are
not perceptually equal, so generating a palette of colors cannot
proceed in a rational manner.  This leads us on to consideration 
to other colorspaces that are derived from CIEXYZ.

\subsection{Perceptual uniformity:  CIELUV}

The CIELUV colorspace
 is a transformation of CIEXYZ in a 
reasonably successful attempt to achieve perceptual uniformity.
Again, any color can be specified as a triplet of values,
with the advantage that a unit step in any direction corresponds to a 
fixed perceptual change in color.  However, the problem of 
intuitively specifying colors remains.

\subsection{Intuitive color choice:  CIELCH}

The CIELCH colorspace is a re-expression of CIELUV using
polar coordinates for the U-V plane.  In this colorspace,
any color may be described by a triplet of values, a unit
step is perceptually uniform throughout the colorspace, 
and the three dimensions of the colorspace have intuitive 
interpretations:

@@

Only one problem remains:  we can conveniently specify a color
using CIELCH, but in order to render that color we need a
specification that a rendering device understands.
For example, in order to display a color on screen,
we need to be able to convert a CIELCH color into an RGB
color specification.

@@

The conversion from CIELCH to RGB consists of a series
of steps.  We start with conversions between the various
CIE colorspaces.

@@

This conversion is non-linear, reflecting the non-linear
nature of human color perception.  It is also performed in 
several stages, reflecting the historical development of the
CIELUV colorspace, which builds upon a couple of other
earlier transformations of the CIEXYZ colorspace.

The conversion also depends on the concept of a \emph{white point}.
For now, we can just think of the white point as representing the
viewing conditions under which the colour is to be
shown.\footnote{It is relatively straightforward 
to think about a white point in terms of a printed image.  In that case,
the white point is an area of the image that perfectly reflects the
ambient lighting conditions.  In other words, the white point
corresponds to whatever illumination is being used to view the image.}
The white point is important because the perception of colors
is dependent on the viewing conditions.  For example, an image
will look different when viewed under a pink light instead
of a white light.  More subtly, 
an image will look different when viewed outside in normal sunlight compared
to inside with fluorescent lighting.

The white point is expressed as CIEXYZ tristimulus values.
There are a number of established white points used in practice,
with the most popular being a representation of normal daylight,
called $D65$:

@@

One such standard is the ITU-R BT.709-2 Recommendation (for
high-definition television), which specifies CIEXYZ values
for the $R$, $G$, and $B$ primaries of a display device.
